Exploring Zstandard user-provided dictionary compression for FASTA files

### 2022-12-10 paper/Persico2022.qmd
Final paper, final code

---
title: "Exploring Zstandard user-provided dictionary compression for FASTA files"
format:
  plos-pdf:
    number-sections: false
    journal:
      # This is an identifier for the target journal: 
      # from https://plos.org/resources/writing-center/ following submission guidelines link, the identifier is the part of the URL after https://journals.plos.org/<id>/s/submission-guidelines
      id: plosbiology
    include-in-header: |
      % Remove comment for double spacing
      % \usepackage{setspace} 
      % \doublespacing
author:
  - name: Michael A. Persico
    equal-contributor: false
    affiliations:
      - ref: aff1
    notes: |
      "Current Address: Dept/Program/Center, Institution Name, City, State, Country"
    corresponding: false
    email: correspondingauthor@institute.edu
author-notes:
  equal-contributor: These authors contributed equally to this work.
  deceased: Deceased
  group: Membership list can be found in the Acknowledgments sections
affiliations:
  - id: aff1
    name: Department of Biology, Concordia University
    city: Montreal
    state: Quebec
    country: Canada
abstract: |
  ## Background

  Zstandard (Zstd) represents a universal, lossless data compression standard and implementation that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance. Previous studies have paired specific Zstd configurations with various file formats in bioinformatics to reduce total data volume. This paper presents a primitive "training mode" model, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements and compare them to the compression performance of Xz, Zlib, Bzip2, and Lz4 universal compressors.

  ## Results

  ## Conclusions
bibliography: bibliography.bib 
---

@@

@@

![Average FASTA formatted data size (bytes) following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd compression. *Escherichia coli* MS 200-1 (NZ_ADUC00000000) artifact for Zstd dictionary generation. Randomized DNA sequences within the range of 100 000 to 150 000 base pairs.](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/Compression_Performance_Boxplot.png){#fig-performance}

@@

The provided FASTA training dataset was retrieved from a set of \textit{E.coli} MS 200-1 strain genomic scaffolds from a whole genome shotgun run[@ncbi]. Sampling a single organism may allow for additional compression/decompression improvements as the sequence data could possess repeating patterns unique to that individual's or species' genome that are recognized and added to the Zstd dictionary during training. The FASTA files themselves may bear similarity with one another in the same set, such as each file being of similar length. Simultaneously, the Zstd dictionary may not achieve comparable compression/decompression performance when attempted on more dissimilar sequence data. Portability of Zstd dictionaries, in the sense of being universally applicable across all sequence data, could be weighed against generating dedicated dictionaries for specific datasets depending on the storage requirements.  Zstd parameters were kept default, though they can be manually altered, including advanced compression options like compression job size or selecting from 22 predefined compression levels[@facebook].

Zstd's multifarious design warrants further research in how tuning the implementation by altering available options (block size, compression level,...) that might influence compression performance. The development of novel data toolsets have utilized Zstd in unique ways. Aside from the Nucleotide Archival Format, FASTAFS was recently introduced as a filesystem in userspace (FUSE) toolkit wherein FASTA files are pooled into a virtualization abstraction layer that archives the set of data synced with its metadata, generating a high-level interface for accessing and storing FASTA data[@fastafs]. The zstd-seekable compression library is an embedded component that keeps the FASTAFS archive compressed for storage reduction purposes, and it represents an alternative compressed data format that deviates from Zstandard, albeit in a compatible manner, that allows for partial decompression of subranges of data[@facebook]. This is achieved via independently compressed/decompressed Zstd frames, which FASTAFS exploits for fast random access to FASTA files by decompressing only data of interest. Further areas of research for general lossless data compression in the context of biological data could include exploring other properties like compression/decompression speed performance, meaning how fast data transcoding occurs, or piloting proof-of-concept algorithms that attempt novel strategies.

# Supplementary information

::: {.supp}
## S1 Fig. {#s1-fig}

Bold the title sentence.

Add descriptive text after the title of the item (optional).
:::

::: {.supp}
## S2 Fig. {#s2-fig}

Lorem ipsum.

Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.
:::

@@

I would like to express my sincere gratitude to Professor David Walsh for teaching the bioinformatics course; My friends and family that supported me throughout my studies; The Julia community for their support before and throughout the writing of this paper; Mark Kittisopikul for developing the CodecZstd.jl #dictionary_and_parameters branch for Zstd dictionary functions; The Quarto developers for producing the Quarto publishing system[@Allaire_Quarto_2022] along with the PLOS template used for this paper; Luca Di Maio and contributors to the Distrobox tool for the ease of setting up the containerized development environments[@maio]; Maximiliano Sandoval and contributors to the Citations app for managing the paper's bibliography[@sandoval], and all other persons that had indirectly assisted through their programs and research.

>>>>

---
title: "Exploring Zstandard user-provided dictionary compression for FASTA files"
format:
  plos-pdf:
    number-sections: false
    journal:
      # This is an identifier for the target journal: 
      # from https://plos.org/resources/writing-center/ following submission guidelines link, the identifier is the part of the URL after https://journals.plos.org/<id>/s/submission-guidelines
      id: plosbiology
    include-in-header: |
      % Remove comment for double spacing
      % \usepackage{setspace} 
      % \doublespacing
author:
  - name: Michael A. Persico
    equal-contributor: false
    affiliations:
      - ref: aff1
    notes: |
      "Current Address: Dept/Program/Center, Institution Name, City, State, Country"
    corresponding: false
    email: correspondingauthor@institute.edu
author-notes:
  equal-contributor: These authors contributed equally to this work.
  deceased: Deceased
  group: Membership list can be found in the Acknowledgments sections
affiliations:
  - id: aff1
    name: Department of Biology, Concordia University
    city: Montreal
    state: Quebec
    country: Canada
abstract: |
  Zstandard (Zstd) represents a universal, lossless data compression standard and implementation that is highly configurable and is aimed at coupling high compression ratios with fast compression/decompression performance. Previous studies have paired specific Zstd configurations with various file formats in bioinformatics to reduce total data volume. This paper presents a primitive "training mode" model, written in the Julia programming language, wherein a custom compression dictionary is generated from a sample FASTA set in order to explore further compression improvements and compare them to the compression performance of Xz, Zlib, Bzip2, and Lz4 universal compressors. Zstd dictionary-based compression did not yield significant compression gains (P < 0.05 in many compressor contrast scenarios using Tukey's HSD(honestly significant difference) test) and had performed worse than alternative compressor types. As an *E.coli* training dataset was used to generate a dictionary to compress randomized data, and with the abundance of options available for manually tuning Zstd's parameters, further compression improvements may be possible.

bibliography: bibliography.bib 
---

@@

\newpage

@@

![Average FASTA formatted data size, in bytes, following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd compression. *Escherichia coli* MS 200-1 (NZ_ADUC00000000) artifact for Zstd dictionary generation[23]. Randomized DNA sequences within the range between 100 000 to 150 000 base pairs.](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/Compression_Performance_Boxplot.png){#fig-performance}

![Tukey's HSD test for determining statistical significance between the compression performance, based on compressed data size, of Bzip2, Xz, Zlib, Zstd with default parameters, and Zstd with a custom dictionary based on the E.coli training dataset[23].](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/Tukeytable.png){#fig-tukey}

\newpage

The Tukey's HSD test paints a picture of the statistical significance of the means of the compressed set of FASTA formatted data between each compressor type. As expected, uncompressed data size and compressed data size differ significantly, with no compressed FASTA file above more than 60 kilobytes based on @fig-performance. With dictionary-based compression applied, Zstd performs similarly to Zstd compression with default parameters (P > 0.05) and, based on the trends of the boxplot and the Tukey's test results, in almost every instance aside from default Zstd compression, dictionary-based compression performed worse than the competing compressor types.

@@

As stated in the Results section, Zstd dictionary-based compression results were surprising in the sense that, at least with the configuration (particular *E.coli* dataset, specific number of randomized FASTA files, completely randomized DNA, RNA, or amino acid sequences), it performed either on par, as was the case with default Zstd compression, or worse than the alternative compressors. The provided FASTA training dataset was retrieved from a set of \textit{E.coli} MS 200-1 strain genomic scaffolds from a whole genome shotgun run[@ncbi]. Sampling a single organism may allow for additional compression/decompression improvements as the sequence data could possess repeating patterns unique to that individual's or species' genome that are recognized and added to the Zstd dictionary during training. The FASTA files themselves may bear similarity with one another in the same set, such as each file being of similar length. Simultaneously, the Zstd dictionary may not achieve comparable compression/decompression performance when attempted on more dissimilar sequence data, which might explain the observed results as randomized sequences were generated for compression performance analysis. Portability of Zstd dictionaries, in the sense of being universally applicable across all sequence data, could be weighed against generating dedicated dictionaries for specific datasets depending on the storage requirements. All other Zstd parameters were kept default, though they can be manually altered, including advanced compression options like compression job size or selecting from 22 predefined compression levels[@facebook].

Zstd's multifarious design warrants further research in how tuning the implementation by altering available options (block size, compression level,...) might yield more significant compression performance. The development of novel data toolsets have utilized Zstd in unique ways. Aside from the Nucleotide Archival Format, FASTAFS was recently introduced as a filesystem in userspace (FUSE) toolkit wherein FASTA files are pooled into a virtualization abstraction layer that archives the set of data synced with its metadata, generating a high-level interface for accessing and storing FASTA data[@fastafs]. The zstd-seekable compression library is an embedded component that keeps the FASTAFS archive compressed for storage reduction purposes, and it represents an alternative compressed data format that deviates from Zstandard, albeit in a compatible manner, that allows for partial decompression of subranges of data[@facebook]. This is achieved via independently compressed/decompressed Zstd frames, which FASTAFS exploits for fast random access to FASTA files by decompressing only data of interest. Further areas of research for general lossless data compression in the context of biological data could include exploring other properties like compression/decompression speed performance, meaning how fast data transcoding occurs, or exploring proof-of-concept algorithms that attempt novel strategies.

\newpage

# Supplementary Material

![Average FASTA formatted data size, in bytes, following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd compression. *Escherichia coli* MS 200-1 (NZ_ADUC00000000) artifact for Zstd dictionary generation. In this instance, randomized DNA sequences were generated within a shorter range between 100 000 to 150 000 base pairs. The discrepancy between uncompressed and compressed data is less pronounced than in @fig-performance, though no compressed FASTA file exceeds more than four kilobytes in size.](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/SupplementaryPlotOne.png){#fig-rna}

![Average FASTA formatted data size, in bytes, following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd compression. *Escherichia coli* MS 200-1 (NZ_ADUC00000000) artifact for Zstd dictionary generation. In this instance, randomized RNA sequences were generated within the range between 100 000 to 150 000 base pairs. The results visually remain similar to that of @fig-performance, although the compression performance of dictionary-based Zstd compression appears to have slightly regressed, perhaps due to some improvements in data compression gained via dictionary use no longer being possible due the difference in sequence type (substitution of thymine with uracil).](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/SupplementaryPlotTwo.png){#fig-rna}

![Average FASTA formatted data size, in bytes, following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd compression. *Escherichia coli* MS 200-1 (NZ_ADUC00000000) artifact for Zstd dictionary generation. In this instance, randomized amino acid sequences were generated within the range between 100 000 to 150 000 codons. Zstd default and dictionary-based compression performance remain similar, it being assumed that the more complex sequence type (20 possible codons versus four possible nucleotides per position) may have erased dictionary compression improvements trained from a DNA sequence dataset.](/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/SupplementaryPlotThree.png){#fig-rna}

\newpage

@@

I would like to express my sincere gratitude to Professor David Walsh for aiding me throughout the project and for teaching the bioinformatics course; My friends and family that supported me throughout my studies; The Julia community for their support before and throughout the writing of this paper; Mark Kittisopikul for developing the CodecZstd.jl #dictionary_and_parameters branch for Zstd dictionary functions; The Quarto developers for producing the Quarto publishing system[@Allaire_Quarto_2022] along with the PLOS template used for this paper; Luca Di Maio and contributors to the Distrobox tool for the ease of setting up the containerized development environments[@maio]; Maximiliano Sandoval and contributors to the Citations app for managing the paper's bibliography[@sandoval], and all other persons that had indirectly assisted through their programs and research.

### 2022-12-10 paper/Persico2022.tex
Final paper, final code

\section*{Abstract}
\hypertarget{background}{%
\subsection{Background}\label{background}}

Zstandard (Zstd) represents a universal, lossless data compression
standard and implementation that is highly configurable and is aimed at
coupling high compression ratios with fast compression/decompression
performance. Previous studies have paired specific Zstd configurations
with various file formats in bioinformatics to reduce total data volume.
This paper presents a primitive ``training mode'' model, written in the
Julia programming language, wherein a custom compression dictionary is
generated from a sample FASTA set in order to explore further
compression improvements and compare them to the compression performance
of Xz, Zlib, Bzip2, and Lz4 universal compressors.

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{conclusions}{%
\subsection{Conclusions}\label{conclusions}}

@@

\caption{\label{fig-performance}Average FASTA formatted data size
(bytes) following Bzip2, Xz, Zlib, default Zstd and dictionary-based
Zstd compression. \emph{Escherichia coli} MS 200-1 (NZ\_ADUC00000000)
artifact for Zstd dictionary generation. Randomized DNA sequences within
the range of 100 000 to 150 000 base pairs.}

@@

@@

The provided FASTA training dataset was retrieved from a set of
\textit{E.coli} MS 200-1 strain genomic scaffolds from a whole genome
shotgun run\citep{ncbi}. Sampling a single organism may allow for
additional compression/decompression improvements as the sequence data
could possess repeating patterns unique to that individual's or species'
genome that are recognized and added to the Zstd dictionary during
training. The FASTA files themselves may bear similarity with one
another in the same set, such as each file being of similar length.
Simultaneously, the Zstd dictionary may not achieve comparable
compression/decompression performance when attempted on more dissimilar
sequence data. Portability of Zstd dictionaries, in the sense of being
universally applicable across all sequence data, could be weighed
against generating dedicated dictionaries for specific datasets
depending on the storage requirements. Zstd parameters were kept
default, though they can be manually altered, including advanced
compression options like compression job size or selecting from 22
predefined compression levels\citep{facebook}.

Zstd's multifarious design warrants further research in how tuning the
implementation by altering available options (block size, compression
level,\ldots) that might influence compression performance. The
development of novel data toolsets have utilized Zstd in unique ways.
Aside from the Nucleotide Archival Format, FASTAFS was recently
introduced as a filesystem in userspace (FUSE) toolkit wherein FASTA
files are pooled into a virtualization abstraction layer that archives
the set of data synced with its metadata, generating a high-level
interface for accessing and storing FASTA data\citep{fastafs}. The
zstd-seekable compression library is an embedded component that keeps
the FASTAFS archive compressed for storage reduction purposes, and it
represents an alternative compressed data format that deviates from
Zstandard, albeit in a compatible manner, that allows for partial
decompression of subranges of data\citep{facebook}. This is achieved via
independently compressed/decompressed Zstd frames, which FASTAFS
exploits for fast random access to FASTA files by decompressing only
data of interest. Further areas of research for general lossless data
compression in the context of biological data could include exploring
other properties like compression/decompression speed performance,
meaning how fast data transcoding occurs, or piloting proof-of-concept
algorithms that attempt novel strategies.

\hypertarget{supplementary-information}{%
\section{Supplementary information}\label{supplementary-information}}

\paragraph*{S1 Fig.}
\label{s1-fig}
{\textbf{Bold the title sentence.}} Add descriptive text after the title
of the item (optional).

\paragraph*{S2 Fig.}
\label{s2-fig}
{\textbf{Lorem ipsum.}} Maecenas convallis mauris sit amet sem ultrices
gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas
ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus
consectetur pellentesque.

@@

I would like to express my sincere gratitude to Professor David Walsh
for teaching the bioinformatics course; My friends and family that
supported me throughout my studies; The Julia community for their
support before and throughout the writing of this paper; Mark
Kittisopikul for developing the CodecZstd.jl
\#dictionary\_and\_parameters branch for Zstd dictionary functions; The
Quarto developers for producing the Quarto publishing
system\citep{Allaire_Quarto_2022} along with the PLOS template used for
this paper; Luca Di Maio and contributors to the Distrobox tool for the
ease of setting up the containerized development
environments\citep{maio}; Maximiliano Sandoval and contributors to the
Citations app for managing the paper's bibliography\citep{sandoval}, and
all other persons that had indirectly assisted through their programs
and research.

>>>>

\section*{Abstract}
Zstandard (Zstd) represents a universal, lossless data compression
standard and implementation that is highly configurable and is aimed at
coupling high compression ratios with fast compression/decompression
performance. Previous studies have paired specific Zstd configurations
with various file formats in bioinformatics to reduce total data volume.
This paper presents a primitive ``training mode'' model, written in the
Julia programming language, wherein a custom compression dictionary is
generated from a sample FASTA set in order to explore further
compression improvements and compare them to the compression performance
of Xz, Zlib, Bzip2, and Lz4 universal compressors. Zstd dictionary-based
compression did not yield significant compression gains (P \textless{}
0.05 in many compressor contrast scenarios using Tukey's HSD(honestly
significant difference) test) and had performed worse than alternative
compressor types. As an \emph{E.coli} training dataset was used to
generate a dictionary to compress randomized data, and with the
abundance of options available for manually tuning Zstd's parameters,
further compression improvements may be possible.

@@

\caption{\label{fig-performance}Average FASTA formatted data size, in
bytes, following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd
compression. \emph{Escherichia coli} MS 200-1 (NZ\_ADUC00000000)
artifact for Zstd dictionary generation{[}23{]}. Randomized DNA
sequences within the range between 100 000 to 150 000 base pairs.}

@@

{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/Tukeytable.png}

\caption{\label{fig-tukey}Tukey's HSD test for determining statistical
significance between the compression performance, based on compressed
data size, of Bzip2, Xz, Zlib, Zstd with default parameters, and Zstd
with a custom dictionary based on the E.coli training dataset{[}23{]}.}

\newpage

The Tukey's HSD test paints a picture of the statistical significance of
the means of the compressed set of FASTA formatted data between each
compressor type. As expected, uncompressed data size and compressed data
size differ significantly, with no compressed FASTA file above more than
60 kilobytes based on Fig~\ref{fig-performance}. With dictionary-based
compression applied, Zstd performs similarly to Zstd compression with
default parameters (P \textgreater{} 0.05) and, based on the trends of
the boxplot and the Tukey's test results, in almost every instance aside
from default Zstd compression, dictionary-based compression performed
worse than the competing compressor types.

@@

As stated in the Results section, Zstd dictionary-based compression
results were surprising in the sense that, at least with the
configuration (particular \emph{E.coli} dataset, specific number of
randomized FASTA files, completely randomized DNA, RNA, or amino acid
sequences), it performed either on par, as was the case with default
Zstd compression, or worse than the alternative compressors. The
provided FASTA training dataset was retrieved from a set of
\textit{E.coli} MS 200-1 strain genomic scaffolds from a whole genome
shotgun run\citep{ncbi}. Sampling a single organism may allow for
additional compression/decompression improvements as the sequence data
could possess repeating patterns unique to that individual's or species'
genome that are recognized and added to the Zstd dictionary during
training. The FASTA files themselves may bear similarity with one
another in the same set, such as each file being of similar length.
Simultaneously, the Zstd dictionary may not achieve comparable
compression/decompression performance when attempted on more dissimilar
sequence data, which might explain the observed results as randomized
sequences were generated for compression performance analysis.
Portability of Zstd dictionaries, in the sense of being universally
applicable across all sequence data, could be weighed against generating
dedicated dictionaries for specific datasets depending on the storage
requirements. All other Zstd parameters were kept default, though they
can be manually altered, including advanced compression options like
compression job size or selecting from 22 predefined compression
levels\citep{facebook}.

Zstd's multifarious design warrants further research in how tuning the
implementation by altering available options (block size, compression
level,\ldots) might yield more significant compression performance. The
development of novel data toolsets have utilized Zstd in unique ways.
Aside from the Nucleotide Archival Format, FASTAFS was recently
introduced as a filesystem in userspace (FUSE) toolkit wherein FASTA
files are pooled into a virtualization abstraction layer that archives
the set of data synced with its metadata, generating a high-level
interface for accessing and storing FASTA data\citep{fastafs}. The
zstd-seekable compression library is an embedded component that keeps
the FASTAFS archive compressed for storage reduction purposes, and it
represents an alternative compressed data format that deviates from
Zstandard, albeit in a compatible manner, that allows for partial
decompression of subranges of data\citep{facebook}. This is achieved via
independently compressed/decompressed Zstd frames, which FASTAFS
exploits for fast random access to FASTA files by decompressing only
data of interest. Further areas of research for general lossless data
compression in the context of biological data could include exploring
other properties like compression/decompression speed performance,
meaning how fast data transcoding occurs, or exploring proof-of-concept
algorithms that attempt novel strategies.

\newpage

\hypertarget{supplementary-material}{%
\section{Supplementary Material}\label{supplementary-material}}

{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/SupplementaryPlotOne.png}

\caption{\label{fig-rna}Average FASTA formatted data size, in bytes,
following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd
compression. \emph{Escherichia coli} MS 200-1 (NZ\_ADUC00000000)
artifact for Zstd dictionary generation. In this instance, randomized
DNA sequences were generated within a shorter range between 100 000 to
150 000 base pairs. The discrepancy between uncompressed and compressed
data is less pronounced than in Fig~\ref{fig-performance}, though no
compressed FASTA file exceeds more than four kilobytes in size.}

{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/SupplementaryPlotTwo.png}

\caption{\label{fig-rna}Average FASTA formatted data size, in bytes,
following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd
compression. \emph{Escherichia coli} MS 200-1 (NZ\_ADUC00000000)
artifact for Zstd dictionary generation. In this instance, randomized
RNA sequences were generated within the range between 100 000 to 150 000
base pairs. The results visually remain similar to that of
Fig~\ref{fig-performance}, although the compression performance of
dictionary-based Zstd compression appears to have slightly regressed,
perhaps due to some improvements in data compression gained via
dictionary use no longer being possible due the difference in sequence
type (substitution of thymine with uracil).}

{\centering \includegraphics{/var/home/mpersico/distrobox/ubuntu-zstd/Persico2022/assets/plots/SupplementaryPlotThree.png}

\caption{\label{fig-rna}Average FASTA formatted data size, in bytes,
following Bzip2, Xz, Zlib, default Zstd and dictionary-based Zstd
compression. \emph{Escherichia coli} MS 200-1 (NZ\_ADUC00000000)
artifact for Zstd dictionary generation. In this instance, randomized
amino acid sequences were generated within the range between 100 000 to
150 000 codons. Zstd default and dictionary-based compression
performance remain similar, it being assumed that the more complex
sequence type (20 possible codons versus four possible nucleotides per
position) may have erased dictionary compression improvements trained
from a DNA sequence dataset.}

\newpage

@@

I would like to express my sincere gratitude to Professor David Walsh
for aiding me throughout the project and for teaching the bioinformatics
course; My friends and family that supported me throughout my studies;
The Julia community for their support before and throughout the writing
of this paper; Mark Kittisopikul for developing the CodecZstd.jl
\#dictionary\_and\_parameters branch for Zstd dictionary functions; The
Quarto developers for producing the Quarto publishing
system\citep{Allaire_Quarto_2022} along with the PLOS template used for
this paper; Luca Di Maio and contributors to the Distrobox tool for the
ease of setting up the containerized development
environments\citep{maio}; Maximiliano Sandoval and contributors to the
Citations app for managing the paper's bibliography\citep{sandoval}, and
all other persons that had indirectly assisted through their programs
and research.
